### 2023-09-27 subfigure/test5.tex

\documentclass{article}

\makeatletter
%% We pretend that we are a naughty class and are \let'ting our
%% \endfigure to the \end@float, etc.

\def\thefigure{\@arabic\c@figure}
\def\fps@figure{tbp}
\def\ftype@figure{1}
\def\ext@figure{lof}
\def\fnum@figure{Fig.~\thefigure}
\def\figure{\@float{figure}}
\let\endfigure\end@float
\@namedef{figure*}{\@dblfloat{figure}}
\@namedef{endfigure*}{\end@dblfloat}

\def\thetable{\@Roman\c@table}
\def\fps@table{tbp}
\def\ftype@table{2}
\def\ext@table{lot}
\def\fnum@table{TABLE~\thetable}
\def\table{%
  \def\@floatboxreset{\reset@font\footnotesize\@setminipage}%
  \@float{table}}
\let\endtable\end@float
\@namedef{table*}{%
  \def\@floatboxreset{\reset@font\footnotesize\@setminipage}%
  \@dblfloat{table}}
\@namedef{endtable*}{\end@dblfloat}
\makeatother

% Next we load the subfigure package and see if it finds the problem.
% To check this test result, you need to see the TTY listing or check
% the test5.log (which we save as test5B.ps).
\usepackage[FIGTOPCAP]{subfigure}
\setcounter{lofdepth}{2}

\newcommand{\figbox}[1]{%
  \fbox{%
    \vbox to 1in{%
    \vfil
    \hbox to 2in{%
      \hfil
      #1%
      \hfil}%
    \vfil}}}

\newcommand{\goodgap}{%
  \hspace{\subfigcapskip}}

\begin{document}
\listoffigures
\newpage
\section{Fifth Test}
This test checks to see if the subfigure package reacts correctly to
the known sort of problem class.  We check that the list-of-figures
page has the two subcaptions in addition to the main caption.  If not,
than the nasty class has zapped us.

(We also need to check the test5.log file to insure that the problem
was identified by the subfigure package and a warning was issued.)

\begin{figure}%
  \centering
  \caption{Figure One.}%
  \label{fig:first}%
  \subfigure[One subone.\label{fig:first:A}]{\figbox{One subone}}%
  \goodgap
  \subfigure[One subtwo.\label{fig:first:B}]{\figbox{One subtwo}}%
\end{figure}

\end{document}